---
format:
  html:
    theme: flatly
    toc: true
    toc-depth: 2
    code-fold: true
    code-tools: false
execute:
  echo: false
  warning: false
  message: false
---

# Introduction

This dashboard compares the **actual 2024 congressional election results** with a **hypothetical rerun** under the district map proposed and passed under California's Prop 50. Using SR precinct-level vote data and an area-weighted interpolation approach, we estimate how seat outcomes shift. Using two metrics, Mean-Median and the Efficiency Gap, we estimate how Prop 50 might affect the "fairness" of the 2026 Midterm Election.


## Map Comparison


:::: {.columns}


::: {.column width="50%"}
### Districts before Prop 50


```{r}
library(ggplot)
library(dplyr)
library(sf)
library(tidyverse)

SR24_with_shp |>
    mutate(winner = case_when(DEM > REP ~ "Dem wins",
      REP > DEM ~ "Rep wins",
      TRUE ~ NA)) |>
    ggplot() +
    geom_sf(aes(fill = winner)) +
    scale_fill_manual(values = c("blue","red")) +
    theme_void() +
    labs(fill = "Winner")
```
:::


::: {.column width="50%"}

### Districts after Prop 50


```{r}
ggplot(prop50_mm) +
geom_sf(aes(fill = winner)) +
scale_fill_manual(values = c("blue","red")) +
theme_void()
```
:::


::::


## Gerrymandering Metrics

### Mean-Median Score before Prop 50: 0.0075
### Efficiency Gap before Prop 50: 
    Democratic votes wasted = 4164545 
    Republican votes wasted = 3822497        
    EG = 0.01305917     

### Mean-Median Score after Prop 50: -0.0217274987828582
### Efficiency Gap after Prop 50: 
    Democratic votes wasted = 2089786
    Republican votes wasted = 3953414       
    EG = -0.1541915  


### The difference in wasted votes was 11.8% in favor of Democrats.
---
